---
title: "Workshop 4"
author: "YH"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 10
    fig_height: 4
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Labor and employment

Query data from https://www.ncsl.org/research/labor-and-employment/national-employment-monthly-update.aspx, which contain unemployment data originally from the Bureau of Labor Statistics

```{r}
library(rvest)
library(tidyverse)
url = "https://www.ncsl.org/research/labor-and-employment/national-employment-monthly-update.aspx"
table = read_html(url) %>% html_elements("table") %>% html_table()
df_unemp <- table[[1]] %>%
  rename(year = 1)
df_unemp

"C:\Program Files\Git\usr\bin\gpg.exe"

# Convert it into tidy format and plot data
df_unemp %>% pivot_longer(January:December, names_to = "month", values_to = "unemp_rate") %>%
  mutate(yr_month = ym(paste(year, month, sep = "-"))) %>%
  ggplot(aes(x = yr_month, y = unemp_rate)) +
  geom_line(lwd = 2, color = "indianred4") +
  theme_minimal() +
  labs(title = "US Unemployment rate, 2008 - Jan 2024",
       y = "Percent unemployed", x = "Year")
```

## Household income

As API calls to FRED requires an API key, we need to sign up for an API key first.

+ It's important to keep your API key private.

+ In the code below, my API key is saved as `Sys.getenv("FRED_KEY")`. This is one way to keep the API key secret. You can find out more about it [here](https://daattali.gitbooks.io/stat545-ubc-github-io/content/bit003_api-key-env-var.html).

```{r}
library(httr)
base_url <- "https://api.stlouisfed.org"
endpoint <- "/fred/series/observations"
resource_url <- paste0(base_url, endpoint)
query_params = list(api_key = Sys.getenv("FRED_KEY"),
                    series_id = "MEHOINUSA672N",
                    file_type = "json")
# Request data from the server
res = GET(resource_url, query = query_params)

# Parse JSON
library(jsonlite)
res_list <- content(res, type = "text") %>% fromJSON(flatten = TRUE)
# Convert data into tibble
income <- as_tibble(res_list$observations)

# Plots
income %>%
  mutate(year = year(date), value = as.numeric(value)) %>%
  filter(year >= 1990) %>%
  ggplot(aes(x = year, y = value)) +
  geom_line(lwd = 2, color = "indianred4", alpha = 0.9) +
  theme_minimal() +
  labs(x = "", y = "", title = "Real Median Household Income in the US, 1990 to present")
```

## Realtime Carpark availability

```{r}
# Construct the resource URL
base_url <- "http://datamall2.mytransport.sg"
endpoint <- "/ltaodataservice/CarParkAvailabilityv2"
resource_url <- paste0(base_url, endpoint)
# Make the request and parse the returned data
res <- GET(resource_url, 
           add_headers(AccountKey = Sys.getenv("LTA_KEY"),
                       accept = "application/json"))
res_content <- content(res, as = "text")
res_list <- fromJSON(res_content, flatten = TRUE)
# Extract information from sublist
df_carpark <- as_tibble(res_list$value)
head(df_carpark)
```

